A constructive Knaster--Tarski proof of the uncountability of the reals

### 2018-08-31 paper-macneille-reals.tex
typo

One way to verify the uncountability of the reals is as follows. We first
observe that the usual diagonalization technique shows that the powerset of the
naturals is uncountable. We then show that this powerset is in bijection with
the reals. In constructive mathematics, more precisely the kind of mathematics
which can be carried out in any topos, the first step is still valid, while the
second might fail -- for any of the several possible flavors of the reals
which in constructive mathematics might not coincide. Therefore a different
approach is needed.

This note gives a constructive proof that a specific flavor of the reals,
namely the MacNeille reals, is uncountable. To the best of our knowledge, this
is the first result in that direction. However, it is just a baby step towards
an understanding whether any of the more interesting flavors of the reals --
for instance the Cauchy reals or the Dedekind reals -- can constructively be
shown to be uncountable. The proof presented here is made possible because the
MacNeille reals -- unlike the Cauchy or Dedekind reals -- can constructively be
shown to be order-complete.

>>>>

One way to verify the uncountability of the reals is as follows. We first
observe that the usual diagonalization technique shows that the powerset of the
naturals is uncountable. We then show that this powerset is in bijection with
the reals. In constructive mathematics, more precisely the kind of mathematics
which can be carried out in any topos, the first step is still valid, while the
second might fail -- for any of the several possible flavors of the reals
such as the Cauchy reals, the Dedekind reals or the MacNeille reals (which all
coincide in classical mathematics, but might differ in constructive
mathematics). Therefore a different approach is needed.

This note gives a constructive proof that one of these flavors,
the MacNeille reals, is uncountable. To the best of our knowledge, this
is the first result in that direction. However, it is just a baby step towards
an understanding whether any of the more interesting flavors of the reals can constructively be
shown to be uncountable. The proof presented here is made possible because the
MacNeille reals -- unlike the Cauchy or Dedekind reals -- can constructively be
shown to be order-complete.
